Syntax Is All You Need: A Universal-Language Approach to Mutant Generation

### 2023-03-30 paper/main.tex
Cleaning up some typos

This diversity of tools highlights the growth of widespread interest in mutation
testing.  However, it also demonstrates various ways that bringing mutation testing to
typically polyglot modern projects is cumbersome. Each language (or IR, at
minimum) requires its own underlying mutation and testing framework.
Mutation tools are not trivial to develop; the smallest
single-language tools we examined were nearly 4KLOC, and the mean size
of a mutation testing tool was over 20K non-comment source lines.

@@

%There
%appears to be a significant risk in relying on any mutation tools other
%than the most widely used and well-supported, such as PIT.
A great deal of research effort has been devoted to devising clever
schemes for reducing the computational burden of mutation testing over
the years (e.g., ~\cite{offutt2001mutation,
  untch1993mutation,KaufmanFAKAJ2022}); in practice, however, the recent adoption of mutation
testing arguably relies more on the availability of vast (cloud) computing
resources than any particular approach from the literature, beyond
heuristic restrictions on the mutants used (e.g., one per line, or
only some siimple operators).   On the other hand, almost no
research has addressed perhaps the most central issue of
mutation testing: in the absence of a working mutation testing
tool for the language in which one is developing software, one cannot
obtain the benefits of mutation testing.  The absence of such tools is
due to the high human cost of developing and maintaining mutation
testing tools.  In particular, new languages may have unstable
grammars, and few tools to aid developers in parsing and manipulating
source code.   Ideally, mutation testing would be available for
\emph{any} newly proposed programming language, as soon as that
language has any community of users at all.  In the present
state-of-the-art, however, it may be years before even rudimentary
tools appear, and they may ``disappear'' shortly there-after.  In
fact, the difficulties of parsing code make bytecode or IL-level
mutation a popular recent choice in tool development, despite the advantages of source-based mutation for
users of mutation testing~\cite{CompareSrcBinary}, including the
ability to read and understand mutants (which is often essential if
one aims to add a test to kill a mutant) and to define mutants in
terms of source-level language (or even project/library)-specific
features.  Moreover, even this arguably less desirable bytecode mutation
approach is not available for languages that do not compile to, e.g.,
Java or LLVM bytecode, such as Haskell.

>>>>

This diversity of tools highlights the growth of widespread interest in mutation
testing.  However, it also demonstrates various ways that bringing mutation testing to
typically polyglot modern projects is cumbersome. Each language (or IR, at
minimum) requires its own underlying mutation and testing framework.
Mutation tools are not trivial to develop; the smallest
single-language tools we examined were nearly 4K LOC, and the mean size
of a mutation testing tool was over 20K non-comment source lines.

@@

%There
%appears to be a significant risk in relying on any mutation tools other
%than the most widely used and well-supported, such as PIT.
A great deal of research effort has been devoted to devising clever
schemes for reducing the computational burden of mutation testing over
the years (e.g., ~\cite{offutt2001mutation,
  untch1993mutation,KaufmanFAKAJ2022}); in practice, however, the recent adoption of mutation
testing arguably relies more on the availability of vast (cloud) computing
resources than any particular approach from the literature, beyond
heuristic restrictions on the mutants used (e.g., one per line, or
only some simple operators).   On the other hand, almost no
research has addressed perhaps the most central issue of
mutation testing: in the absence of a working mutation testing
tool for the language in which one is developing software, one cannot
obtain the benefits of mutation testing.  The absence of such tools is
due to the high human cost of developing and maintaining mutation
testing tools.  In particular, new languages may have unstable
grammars, and few tools to aid developers in parsing and manipulating
source code.   Ideally, mutation testing would be available for
\emph{any} newly proposed programming language, as soon as that
language has any community of users at all.  In the present
state-of-the-art, however, it may be years before even rudimentary
tools appear, and they may ``disappear'' shortly there-after.  In
fact, the difficulties of parsing code make bytecode or IL-level
mutation a popular recent choice in tool development, despite the advantages of source-based mutation for
users of mutation testing~\cite{CompareSrcBinary}, including the
ability to read and understand mutants (which is often essential if
one aims to add a test to kill a mutant) and to define mutants in
terms of source-level language (or even project/library-specific
features).  Moreover, even this arguably less desirable bytecode mutation
approach is not available for languages that do not compile to, e.g.,
Java or LLVM bytecode, such as Haskell.
